---
title: "Foundations_of_Stats_Homework"
author: "NYC Data Science Academy"
output: html_document
---
## Question #1: Body Temperature

The `Temp.txt` dataset records the body temperatures, gender, and heart rate of 130 individuals; you may assume that the observations are independent of one another.

1. Read the data into R and provide basic numerical EDA to describe the data.
  + Provide univariate EDA for all variables, and bivariate EDA where appropriate.

2. Provide basic graphical EDA to describe the data.
  + Provide univariate EDA for all continuous variables, and bivariate EDA where appropriate.

3. You have heard that the average human body temperature is 98.6 degrees Fahrenheit. Does this data support this claim? (Perform a hypothesis test to determine whether the true population mean body temperature is 98.6 degrees Fahrenheit.)
  + What is a 95% confidence interval for the average human body temperature?
  + Interpret your results in context of the problem.

4. Is there a significant difference in body temperature between males and females? If so, quantify this difference.
  + What is a 95% confidence interval for the average difference in human body temperature between males and females?
  + Interpret your results in context of the problem.
  + [**Bonus**] If we had collected a lot of samples of the body temperature for different genders, we could have directly compared *the average of the average temperature* for each gender. Hypothesis tests we have covered allow us to deduce the distribution of the average without needing to collect many samples (we have only 65 males and 65 females). Another simple way to achieve this is to **bootstrap**. Random sample (with replacement) 65 observations from each gender, then use them as a new sample. By bootstrapping, derive the 95% confidence interval for the average difference in human body temperature between males and females.

5. You believe the variances of male heart rate and female heart rate are different; specifically, you claim that females have a lower heart rate variance. Test this claim.
  + What is a 95% confidence interval for the ratio between female and male heart rate variances?
  + Interpret your results in context of the problem.

```{r, include=FALSE}
temp = read.table('https://s3.amazonaws.com/graderdata/Temp.txt', header = TRUE)
temp

## part 1.1
summary(temp)
sapply(temp, sd)
## Basic numarical univariate EDA -->
## These observations are equally distributed across Gender, means Female and male both has 65 counts of observations. standard deviation will be 0.5.
## Body temperature min is 96.30 and max is 100.80, while mean is 98.25. while here std is 0.733.
## Heart.Rate min is 57.00 and max is 89.00 while mean is 73.76. std of heart rate is 7.06.

plot(density(temp$Body.Temp), main = "Distribution of Body temp") +
abline(v = mean(temp$Body.Temp), lwd = 2, lty = 2)

plot(density(temp$Heart.Rate), main = "Distribution of Heart Rate") +
abline(v = mean(temp$Heart.Rate), lwd = 2, lty = 2)

## part 1.2

boxplot(temp$Body.Temp[temp$Gender=='Male'], temp$Body.Temp[temp$Gender=='Female'], main ="Distribution of body temperature dataset", col = c("red", "blue"), names = c("Male", "Female"))

## by analyzing boxplot for body temp, it looks like on average female has slightly higher body temp than male.

boxplot(temp$Heart.Rate[temp$Gender=='Male'], temp$Heart.Rate[temp$Gender=='Female'], main ="Distribution of heart rate dataset", col = c("red", "blue"), names = c("Male", "Female"))

## by analyzing box plot for heart rate, looks like female has higher heart rate than male.


## part 1.3 - hypothesis test using t.test function 
## this is a one-sample test
t.test(temp$Body.Temp, mu = 98.6, alternative = "greater")

## The p-value for this test is 1; thus NULL hypothesis is TRUE and this dataset also confirms that average body temp is 98.6 in favor of NULL hypothesis.
## 95 percentage confidance interval also suggest that 98.14 times it can be found that average body temp is 98.6.

## part 1.4 
## To understand the significant difference in body temperature between males and females, we need to conduct One-Way ANOVA test:

boxplot(temp$Body.Temp[temp$Gender=='Male'], temp$Body.Temp[temp$Gender=='Female'], main ="Distribution of body temperature dataset", col = c("red", "blue"), names = c("Male", "Female"))

summary(aov(temp$Body.Temp ~ temp$Gender))

## p-value for this test is 0.02 which is less than our threshold of 0.05.
#Thus, we conclude in favor of the alternative hypothesis; our data supports the
#notion that the true average body temp for Male and Female is different.

# We are specifically interested in the quantification of this body temperature difference between Male and Female. Hence, we conduct F-test to determine the variance:

var.test(temp$Body.Temp[temp$Gender=='Male'],
         temp$Body.Temp[temp$Gender=='Female'],
         alternative = "two.sided")

#The p-value for this test is 0.6211 meaning that, under the null hypothesis, we
#would expect to see results at least as extreme as about 62.11% of the time.
#Thus, our data does not support the alternative hypothesis that the variance among
#body temperature is different for male and female. Hence, p-value clearly supports NULL hypothesis, which means ratio of variance is 1 and hence body temperature is same.
## 95% confidance interval is 53.87%


## part 1.5

var.test(temp$Heart.Rate[temp$Gender=='Male'],
         temp$Heart.Rate[temp$Gender=='Female'],
         alternative = "two.sided")

#The p-value for this test is 0.011 which is less than our threshold of 0.05.
#Thus, we conclude in favor of the alternative hypothesis; male heart rate and female heart rate are different.
## 95% confidance interval is also suggest that this happens between 32.04 to 86.14 times overall.

var(temp$Heart.Rate[temp$Gender=='Male']) ## this gives variance of 34.51 
sd(temp$Heart.Rate[temp$Gender=='Male']) ## this gives std of 5.8
var(temp$Heart.Rate[temp$Gender=='Female']) ## this gives variance of 65.69
sd(temp$Heart.Rate[temp$Gender=='Female']) ## this gives std of 8.10

## Thus it is clear that female has higher variance in heart rate then male and our claim for this is not correct.

```

## Question #2: Plant Growth

Load the `PlantGrowth` dataset located in the `datasets` library; this dataset contains dried plant weight measurements for the same species of plant under three different conditions (two separate growth treatments, and a control group where no treatment was applied). You may assume that the observations are independent of one another.

1. Create side-by-side boxplots of the plant weights segmented by the type of applied treatment. Describe the features of the graph.

2. Calculate the standard deviations of each conditional distribution of plant weight based on the applied treatment. Do these differ significantly?
  + **NB**: To avoid increasing our chance of encountering a "false positive," we must avoid applying three separate F-tests (treatment #1 vs treatment #2; treatment #1 vs control; treatment #2 vs control). As an alternative, Bartlett's Test of Homogeneity of Variances allows us to simultaneously test for the similarity of a group of variances, rather than just a pair. Implement this test using the `bartlett.test()` function and report your results.
  
3. Is there a significant difference in the weight of plants based on the growth treatments they were given? Conduct a hypothesis test and report your results in context of the problem. 
  + Given the results of the Bartlett test, is the result of your hypothesis test valid?



```{r}
library(datasets) #Load the base R datasets library.

help(PlantGrowth) #See the help documentation for the PlantGrowth dataset to see a
               #description of the included variables.

PlantGrowth

## part 2.1
boxplot(PlantGrowth$weight ~ PlantGrowth$group,
        col = c("red", "orange", "yellow"),
        main = "Distribution of PlantGrowth\nfor Various Treatment Types")

## part 2.2
sd(PlantGrowth$weight[PlantGrowth$group=='ctrl']) ## this returns 0.58 
sd(PlantGrowth$weight[PlantGrowth$group=='trt1']) ## this returns 0.79
sd(PlantGrowth$weight[PlantGrowth$group=='trt2']) ## this returns 0.44

## As we can see by standard deviation is very small.

bartlett.test(weight ~ group, data = PlantGrowth)

## p-value for this test is 0.23 which is greater than our threshold of 0.05.
#Thus, we conclude in favor of null hypothesis; meaning the variances in each of the groups (samples) are the same.

## part 2.3
chisq.test(as.table(PlantGrowth$weight))

## p-value is greater than 0.05, which support null hypothesis. Hence, there is not significant difference in weight of plants based on the growth treatments they were given

```

## Question #3: Gender, Hair, & Eye Color

Load the `HairEyeColor` dataset located in the `datasets` library; this is a three dimensional dataset that records the hair color, eye color, and gender of 592 different statistics students.

1. Visualize the entire dataset using a mosaic plot using the following command:
  + `mosaicplot(HairEyeColor, shade = TRUE)`
  + Which category combinations receive more observations than expected? Fewer observations than expected?
  
2. Reduce your dataset to focus on just females with brown and blue eyes (but still include all hair colors). Create another mosaic plot and describe what you see.
  + Conduct a hypothesis test to see if hair and eye color are independent of one another for this reduced dataset. Report your results in context of the problem.

3. For the reduced dataset, which category combination contributed most to any statistical deviations? Which category contributed the least? By how much for each?


```{r}
library(datasets) #Load the base R datasets library.

help(HairEyeColor) #See the help documentation for the HairEyeColor dataset to see a
               #description of the included variables.

HairEyeColor

mosaicplot(HairEyeColor, shade = TRUE)


```

