---
title: "Note on Random Variables"
author: ""
date: "10/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
```

## What a Random Variable is about?

A **random variable** is a variable, whose value is to be determined with a **distribution**. A distribution tells you the likelihood of a possible value (event) to occur.

- A statement like "flipping a fair coin" describes a distribution. It means that the possible events are **head** or **tail**, and the likelihood (probability) for each to occur is 50%. This is Bernoulli distribution. 

- Suppose a random variable has a normal distribution (often denoted by $X \tilde{} N(\mu, \sigma)$), if the possible values are real numbers and the likelihood (probability density) for each number to occur can be represented by the famous *bell curve*.


```{r}
pts = -3 + (1:600)/100
ggplot() +
  geom_line(aes(x=pts, y=dnorm(pts)))
```


## Probability Density

Probability density are not the same as probability, but it does indicates the "likelihood" for a value to occur:

```{r cars}
data = rnorm(100000)

ggplot() + geom_histogram(
  aes(x=data), color="red", fill="yellow"
) + theme_bw()
```

## Side Issue: Why Density

For the same histogram, change the parameter `bins` and inspect the y axis:

```{r}
ggplot() + geom_histogram(
  aes(x=data), bins = 10, color="red", fill="yellow"
) + theme_bw()
```


**Density** doesn't have the same problem:

```{r}
ggplot() + geom_histogram(
  aes(x=data, y= ..density.. ), 
  bins = 10, color="red", fill="yellow"
) + theme_bw()
```


## Other Numerical Distribution

**Unif distribution**
```{r}
data = runif(100000)

ggplot() + geom_histogram(
  aes(x=data), bins = 50, color="red", fill="yellow"
) + theme_bw()
```

**ChiSq distribution**
```{r}
data = rchisq(100000, df = 20)

ggplot() + geom_histogram(
  aes(x=data), bins = 50, color="red", fill="yellow"
) + theme_bw()
```

**F distribution**
```{r}
data = rf(100000, df1 = 50, df2=40)

ggplot() + geom_histogram(
  aes(x=data), bins = 50, color="red", fill="yellow"
) + theme_bw()
```


## Transformations of Distributions

If we know the distribution of a random variable, can we deduce the distribution of a new distribution "transformed" from the old ones? For example, if we know the distribution of a uniform distribution $U$, what about $U+2$?

```{r}
data = runif(100000)

ggplot() + 
  geom_histogram(aes(x=data), bins=50, fill="yellow", alpha= 0.7) + 
  geom_histogram(aes(x=data + 2), bins=50, fill="green", alpha= 0.7)
```


## Central Limit Theorem

The **central limit theorem** is the fundation of many statistical techniques. It's also about deducing the distribution of a new derived random variable from that of the old ones.

Consider the case, that $U_1, U_2, U_3, ..., U_n$ are `n` random variables that are
- mutually independent
- with uniform distribution

What would the distribution of the average, $ \sum_{1}^n(U_i)/n$ be?

Central limit theorem tells us that the average would (approximately) follow the normal distribution with the mean equals to the mean of $U_i$, and with the variance equal to the variance of $U_i$ divided by n.

Let's look at the code to confirm it:

```{r}
n = 300
sample_size = 10000

data = as.data.frame(
  sapply(1:sample_size, function(x){runif(n)})
)

data[, 1:4]
```

so the average of each sample set is:

```{r}
avgs = colSums(data)/n
avgs[1:4]
```

Therefore the distribution can be visualized with the histogram

```{r}
ggplot() +
  geom_histogram(aes(x=avgs))
```

The central limit theorem says not only the distribution is normal, it also specifies the mean and the variance of the normal distribution:

```{r}
mu = 0.5
var = 1/12
ref = rnorm(sample_size, mean=mu, sd=sqrt(var/n))

df = data.frame(
  value = c(avgs, ref), 
  group=rep(c("simu", "ref"), each=sample_size)
)
ggplot(data=df) + 
  geom_histogram(aes(x=value, fill=group), bins=30, alpha= 0.5) +
  facet_grid(group~.)
```

## Another example for central limit theorem

In fact, the central limit theorem is valid for not only a collection of uniform distribution, but **any** distribution:

```{r}
n = 300
sample_size = 10000

dgf = 10
data = as.data.frame(
  sapply(1:sample_size, function(x){rchisq(n, df=dgf)})
)

data[, 1:5]
```

```{r}
hist(as.numeric(data[1,]))
```

then we compute the average in the same way:

```{r}
avgs = colSums(data)/n
avgs[1:5]
```

Again we compare the histogram of our simulation with an actual normal distribution:

```{r}
mu = dgf
var = 2*dgf
ref = rnorm(sample_size, mean=mu, sd=sqrt(var/n))

df = data.frame(
  value = c(avgs, ref), 
  group=rep(c("simu", "ref"), each=sample_size)
)
ggplot(data=df) + 
  geom_histogram(aes(x=value, fill=group), bins=30, alpha= 0.5) +
  facet_grid(group~.)
```

# Still another example for central limit theorem

Now let's try f distribution:

```{r}
n = 300
sample_size = 10000

dgf1=30; dgf2=20
data = as.data.frame(
  sapply(1:sample_size, function(x){rf(n, dgf1, dgf2)})
)

data[, 1:5]
```

```{r}
hist(as.numeric(data[1,]))
```

Finally we compare the histogram:
```{r}
avgs = colSums(data)/n

mu = dgf2/(dgf2 - 2)
var = 2*dgf2^2*(dgf1+dgf2-2)/dgf1/(dgf2-2)^2/(dgf2-4)
ref = rnorm(sample_size, mean=mu, sd=sqrt(var/n))

df = data.frame(
  value = c(avgs, ref), 
  group=rep(c("simu", "ref"), each=sample_size)
)
ggplot(data=df) + 
  geom_histogram(aes(x=value, fill=group), bins=100, alpha= 0.5) +
  facet_grid(group~.)
```


## More Transformations

The  chi square distribution and the f distribution are defined by the transformations.

## Chi Square Distribution

The **chi square distribution with degree of freedom n** is defined to be the distribution of the sum of the square of `n` mutually independent random variables with the standard normal distribution. 

Equivalently, suppose $U_1, U_2, U_3, ..., U_n$ are random variables with the standard normal distribution, then the **chi square distribution with degree of freedom n** is defined to be the distribution of $\sum_i^n U_i^{2}$. (Why is this sum interesting?)

To make it precise:

```{r}
n = 30
sample_size = 10000

data = as.data.frame(
  sapply(1:sample_size, function(x){ rnorm(n)  })
)
data[, 1:5]
```

```{r}
hist(as.numeric(data[1,]))
```

```{r}
SqSum = colSums(data*data)
SqSum[1:4]
```

```{r}
ref = rchisq(sample_size, df = n)

df = data.frame(
  value = c(SqSum, ref), 
  group=rep(c("simu", "ref"), each=sample_size)
)

ggplot(data=df) + 
  geom_histogram(aes(x=value, fill=group), bins=30, alpha= 0.5) +
  facet_grid(group~.)
```

## F Distribution

The **f distribution** is defined to be the distribution of the weighted ratio of two random variables of chi square distribution. 

More precisely, suppose $\chi_{n_1}$ and $\chi_{n_2}$ are, respectively, the chi square distribution of degree of freedom $n_1$ and the chi square distribution of degree of freedom $n_2$. The the f distribution $F_{n_1, n_2}$ is defined to be the distribution of $\frac{ \chi_{n_1}/n_1  }{  \chi_{n2}/n_2 }$.

To simulate:

```{r}
n1=20; n2=30
sample_size = 20000

X1 = sapply(1:sample_size, function(x){ rchisq(1, n1)  })

X2 = sapply(1:sample_size, function(x){ rchisq(1, n2)  })

X1[1:5]
X2[1:5]
```

```{r}
hist(X1)
```


```{r}
ratio = (X1/n1) / (X2/n2)

ref = rf(sample_size, n1, n2)

df = data.frame(
  value = c(ratio, ref), 
  group=rep(c("simu", "ref"), each=sample_size)
)

ggplot(data=df) + 
  geom_histogram(aes(x=value, fill=group), bins=30, alpha= 0.5) +
  facet_grid(group~.)
```






