---
title: "Simple_Linear_Regression_Homework"
author: "NYC Data Science Academy"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
```

## Question 1: Anatomical Data from Domestic Cats
**Purpose: Demonstrating understanding of how to run the models**

1. **Load: ** 
    + Load the *cats* from the **MASS** library. 
    + This dataset includes the body and heart weights of both female and male adult domestic cats. 
```{r}
# Your code here

```

2. **Visualize: **
    + Create a scatterplot of heart weight versus body weight. 
    + From this plot alone, do you think simple linear regression would be a good fit for the data? Explain.

```{r}
# Your code here

```

3. **Model: ** Regress heart weight onto body weight. For this model:
      + Write out the regression equation.
      + Interpret the meanings of the coefficients in context of the problem.
      + Are the coefficients significant? How can you tell?
      + Is the overall regression significant? How can you tell? How does your answer from Part C relate?
      + Find and interpret the RSE?
      + Find and interpret the coefficient of determination.
    
```{r}
# Your code here

```

## Question 2: Machine Learning Theory
**Purpose: Demonstrate theory of lecture material**

1. **Assumptions: **
    + Assess each of the assumptions of linearity. 
    + Which assumption are you tring to correct if you decide to apply a Box-Cox transformation?

```{r}
# Your code here

```


2. **Box-Cox Transformation: ** 
    + If you apply a Box-Cox transformation to a variable which you already applied a Box-Cox transformation, what would happen? 
    + If you are unsure, try doing it to see!

```{r}
# Your code here

```


3. **Evaluation: ** 
    + Why do we use the R^{2} coefficient of determination to validate a linear model rather than the RSS? 
    + Why does it make sense to square the residuals to help determine the best model?

```{text}
# Your explanation here

```


4. **Prediction and Confidence Bands: ** Why is the prediction band wider than the confidence band? Why does the confidence band widen as it travels away from the center of the regression line?

```{text}
# Your explanation here

```

## Question 3: Challenge Questions
**Purpose: Push yourself for more advanced topics**

These questions are an extension of the Question 1 coding questions from above.

1. **Visualize: **Add to your plot from Question 1 Part 2. 
    + Add the regression line.
    + Add the residuals. Do any of the residuals seem abnormally large?
    + Construct 95% confidence intervals for the model coefficients. Interpret the intervals in context of the problem.

```{r}
# Your code here

```

2. **Apply: **Construct confidence and prediction intervals for body weights of 2.8 kg, 5 kg, and 10 kg. Do you foresee any issues with reporting any of these intervals?

```{r}
# Your code here

```

3. **Tune: **Transform your data using a Box-Cox transformation. 
    + Create a Box-Cox plot.
    + Choose the best value of lambda (Keep in mind interpretability).
    + Create a new regression and interpret your results.

```{r}
# Your code here

```

4. **Visualize: **Plot the regression line from the Box-Cox model on the scatter plot of heart weight versus body weight.

```{r}
# Your code here

```

5. **Compare: **Compare the models you created:
    + Give one reason why you might use the original model instead of the Box-Cox transformed model.
    + Give one reason why you might use the Box-Cox transformed model instead of the original model.

```{text}
# Your explanation here

```
