---
title: "Missingness_Imputation_Homework"
author: "NYC Data Science Academy"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
```

## Question #1: Missingness & Imputation for the Titanic Dataset
Load the `titanic3` dataset from the `PASWR` library; this dataset describes the survival status of individual passengers from the Titanic voyage. There are 14 different variables, some of which have quite a bit of missingness.

1. How many variables contain at least one missing value?
    + What are these variables?
    + For each variable, what is the extent of missingness (how many missing values are there and what is the percentage of missingness)?

```{r}
# Your code here

```


2. How many observations contain at least one missing value?
    + What is the percentage of missingness from an observation standpoint? (Hint: the complete.cases() function might be useful here).

```{r}
# Your code here

```


3. How many cells in the data are missing values?
    + What is the percentage of missingness from a dataset standpoint?

```{r}
# Your code here

```


4. What are the different combinations of missingness in the dataset?
```{r}
# Your code here

```


5. What kind of missingness do you have for each variable that contains missing values? Give a reason and scenario as to why you believe this.

```{r}
# Your code here

```

6. Impute using mean value imputation for the age variable.
    + Graph the distributions of the age variable before and after mean value imputation. Describe what you see. What problems may arise?
    
```{r}
# Your code here

```


7. Impute using simple random imputation for the age variable.
    + Graph the distributions of the age variable before and after simple random imputation. Describe what you see. What problems may arise?

```{r}
# Your code here

```


## Question #2: K-Nearest Neighbors with the Titanic Dataset
Continue with the `titanic3` dataset from the `PASWR` library.

1. Impute using the single missing value of the fare variable using simple random imputation. What value was imputed?

```{r}
# Your code here

```


2. Plot the simple random imputation of fare against the simple random imputation of age; color this plot by pclass. Describe any trends.

```{r}
# Your code here

```


3. Add two points to your plot representing the following passengers:
    + A 50 year old who paid $400 for their ticket.
    + A 10 year old whose parents paid $100 for their ticket.
    
```{r}
# Your code here

```


4. What classes would you think these new individuals would belong to?

```{text}
# Your explanation here

```


5. Impute the missing class values for the new passengers using 1 Nearest Neighbor. What were the predicted classes for each passenger?

```{r}
# Your code here

```


6. Impute the missing class values for the new passengers using the $\sqrt{n}$ Nearest Neighbor rule. What were the predicted classes for each passenger? Why did they change/not change?

```{r}
# Your code here

```

## Question #3: Minkowski Distances with the Titanic Dataset
Continue with the `titanic3` dataset from the `PASWR` library.

1. Create a new data frame that includes:
    + The pclass, survived, sex, age, sibsp, and parch variables from the original `titanic3` dataset.
    + The simple random imputation of the fare variable you created above.

```{r}
# Your code here

```


2. Separate this new data frame into two separate data frames as follows (note that there should be no observations that appear in both data frames):
    + For observations that are totally complete: all variables.
    + For observations that are missing a value for age: all variables except age.

```{r}
# Your code here

```


3. Use 1 Nearest Neighbor to impute using:
    + Manhattan distance.
    + Euclidean distance.
    + Minkowski distance with $p = 10$.

```{r}
# Your code here

```


4. Overlay and label four separate density curves: one for each of the three 1 Nearest Neighbor imputed age values, and one for the original complete age observations. Describe what you see any why this might be occurring.

```{r}
# Your code here

```


5. Use the $\sqrt{n}$ Nearest Neighbor rule to impute using:
    + Manhattan distance.
    + Euclidean distance.
    + Minkowski distance with $p = 10$.

```{r}
# Your code here

```


6. Repeat part 4 with the $\sqrt{n}$ Nearest Neighbor solutions. What is happening here?

```{r}
# Your code here

```




